\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-02-18 eg6-letters/NIM/EG6_RTPC_NIM.tex
NIM:  various updates

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
Laboratory (JLab) to track low-energy nuclear recoils for the purpose of 
measuring exclusive nuclear channels, such as coherent Deeply Virtual Compton 
Scattering (DVCS) and coherent meson production on $^4$He. In such 
processes, the $^4$He nucleus remains intact and recoils as a whole in the final 
state. The CEBAF Large Acceptance Spectrometer (CLAS) at JLab has been used for 
similar studies on proton targets, however it cannot track recoils with momenta 
less than 250 MeV/c. In 2009, we carried out measurements using the CLAS spectrometer supplemented by the RTPC positioned directly around a gaseous $^4$He target, allowing a detection threshold as low as $~$100  MeV/c for 
$^4$He. This article discuses the design, work principle, calibration 
methods and performances of the RTPC.
\end{abstract}

@@

Thomas Jefferson National Accelerator Facility (TJNAF),
in Newport News, Virginia, USA, provides 
high power electron beams of energy up to 6 GeV and 100$\%$ duty cycle to 
three experimental Halls (A, B, C) simultaneously. The CLAS spectrometer, located in Hall-B, 
is composed of several sub-detectors and two magnets. The spectrometer was 
designed to track charged particles of momentum greater than 250 MeV/c as well 
as detecting real photons. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) used for the tracking of charged 
    particles~\cite{DCref} (in blue on the figure).
 \item Superconducting toroidal magnet (in yellow) bends the trajectories 
    of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Cherenvkov Counters (CC) used to distinguish electrons from other negative 
    particles, such as pions, at lower energies~\cite{CCref} (in pink).
 \item Scintillation Counters (SC) used to identify hadrons by measuring their 
    time of flight (TOF)~\cite{TOFref} (in red).
 \item Electromagnetic Calorimeters (EC) used to distinguish electrons from other negative 
    particles at higher energies and to detect neutral particles (neutrons and 
    photons in particular)~\cite{ECref} (in green).
\end{itemize}

@@

Measurement of the DVCS process ($eH \rightarrow e' H' \gamma$, where 
$H$ is a hadron) necessitates some adaptations in this setup.
With a 6 GeV electron beam, the 
majority of DVCS photons are produced at very forward angles where the  
acceptance of the EC is poor. To extend the range, an additional 
Inner Calorimeter (IC) was built for the E01-113 experiment in 2005. The IC is 
constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long,
corresponding to 17 radiation lengths. The achieved energy 
resolution is around 3 to 4$\%$ for photon energies between 2 GeV and 5 GeV, with
an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}.

%M{\o}ller electrons can be
%very problematic at such small angles, so a 5 Tesla solenoid magnet has been
%added around the target as a magnetic shield against this low energy background.

For coherent nuclear processes, the RTPC was developed to extend the tracking range of CLAS to low energy nuclei and was used during a one month experimental run in 2009 
with a longitudinally polarized electron beam of energy 6.064 GeV incident 
on a gaseous $^{4}$He target.
%, with the goal of measuring asymmetries in Deep Virtual Compton 
%Scattering (DVCS)~\cite{proposal} and exclusive meson production.

@@

\section{CLAS-EG6 RTPC design} \label{sec_design}
The design of CLAS-EG6 RTPC has been chosen according to physics and 
experimental requirements.  At 6 GeV incident electron beam energy, the recoil 
$^{4}$He nuclei from coherent DVCS have an average momentum (per 
charge) around 125 MeV/c, while the ``standard'' CLAS detects charged particles 
above a threshold of 250 MeV/c. The chosen design also has 
to fit within the available space inside a 5 Tesla Solenoid magnet
surrounding the target \cite{Hyon-suk}. Figure \ref{fig:RTPC2} shows a photograph of the RTPC before being inserted into the solenoid.

The CLAS-EG6 RTPC is a 200~mm long cylinder with a diameter of 150~mm. While a charged 
particle traverses the sensitive gas volume, ionized electrons are released 
and drift, under the effect of an applied electromagnetic field, outward 
perpendicularly to the beam direction. These electrons are then amplified by 
three layers of GEMs and detected by a curved readout system on the external 
shell as can be seen in the figure~\ref{fig:RTPC_1_4}. As for CLAS-BoNuS, the 
EG6-RTPC has two detection halves but unlike BoNuS they are part of a single 
chamber. Each half has independent GEM amplification systems that cover around 
80\% of the azimuthal angle.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of CLAS-EG6 RTPC taken on a plane perpendicular to 
the beam line, see text for description of the elements.} 
\label{fig:RTPC_1_4}
\end{figure}

As seen in figure~\ref{fig:RTPC_1_4}, the CLAS-EG6 RTPC has the following 
substructure, from the beam line to the exterior:\\
\begin{itemize}
   \item The target extends along the RTPC's central z-axis, with a diameter of 
      6 mm. It is enclosed in a 27-$\mu$m-thick Kapton wall.

   \item The first gas gap extends from 3 mm to 20 mm radial distances. It is 
      filled with $^{4}$He gas at 1~atm. During run, this region is swarmed by 
      M\o ller electrons induced by the beam, such that filling this region 
      with a light gas like $^{4}$He minimizes secondary interactions. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and connected to ground.                                                                
   \item The gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode.

   \item The drift region is filled with the Ne-DME gas mixture, it extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is around 500~V/cm 
      (???) and perpendicular to the beam. Field cages are installed at both 
      ends of the detector in this region. Progressive voltage depending on 
      radial position is applied on these to maintain a uniform electric field 
      perpendicular to the beam line at the edges of the RTPC.

   \item The electron amplification system is composed of three GEMs located at 
      radii, 60 mm, 63 mm and 66 mm. In this configuration, the first GEM layer 
      represents the anode.

   \item The readout board has internal radius of 69~mm and collects the 
      charges. Preamplifiers are plugged directly on the outer side and are 
      transmitting the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM system has been mainly chosen for this detector for its flexibility, 
allowing the curved amplification surface. Also GEMs are known to create less 
sparks than in other gaseous detectors, which provides a cleaner detection of 
slow nuclei like $^4$He. GEMs are made from an insulator layer, Kapton in our 
case, sandwiched between two 5 $\mu$m copper layers. The mesh of each GEM layer 
is chemically etched with 50$\mu$m diameter holes in double-conical cross 
section shapes as can be seen in figure~\ref{fig:GEMs}. A potential difference 
of 400 V is applied between the two copper layers of each GEM creating a very 
strong electric field in the holes. Such strong fields lead to high ionization 
from initial electrons and therefore amplification of the signal. Also, a 
potential difference of 150 V is set between the GEMs to push the amplified 
electrons toward the readout pads. The gain of each GEM layer is of order 100, 
making a total gain of about $10^{6}$.\\

\section{Readout System} \label{sec_readout}
The RTPC electron collecting system has 3200 readout pads. These collection 
elements are located at the end of the amplification region, 69 mm from the 
central axis.

@@

Figure \ref{fig:PADs} shows a schematic drawing of the size and the alignment 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width the shift between the rows allows to reduce aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in the 
figure~\ref{fig:PADs} shows how pads are grouped to the 16 channels 
pre-amplifiers. The Time to Digital Converter (TDC) units are 114~ns and 
indicates the time taken by the electron to drift from the ionization point to 
the readout board. On the the other side, the charge information is measured in 
Analog-to-Digital-Converter (ADC) units without specific normalization.

@@

@@

We can measure the drift speed using the tracks detected in the RTPC, in figure 
\ref{fig:RTPC_signals} we represent a typical $^{4}$He track (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads while the geometrical symmetry insure they always travel the same 
distance. So by identifying the maximum TDC measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

We first measure the drift speed along the 200 mm RTPC's length to take into 
account variations in the electric and magnetic field, some of which are known 
(see figure \ref{fig:B_MAP}) some might be due to slight shifts in the 
geometry. We verified this by looking at the TDC profile generated in the 
chamber by all our tracks (see figure \ref{fig:TDC_profile}). The time profile 
of registered hits clearly show the dropping edge expected from the geometrical 
considerations previously mentioned. We define a value named $TDC_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram, this value is inversely proportional to the drift speed. The result 
can be seen in figure \ref{fig:RunNumber_61551_TDCmax_Zslice} showing the 
dependence of drift speed with position along the beam line axis in the 
chamber.

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.35]{fig/B_MAP.png}
   \caption[]{Schematic draw in of the solenoid magnetic vectors inside the 
   RTPC. In this draw, r is the radial distance from the target's axis and z is 
the longitudinal position along the RTPC with its center at zero.} 
\label{fig:B_MAP}
\end{figure}

@@

Due to the non perfect experimental conditions and in particular changes in the 
gas proportions, the drift speed might also change over the 3 months run.  
Figure \ref{fig:Drift_run_number_1} shows the $TDC_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation over time that we parametrized with a fit in order to take it into 
account in the track reconstruction code. Dealing with such situation is 
possible by adapting the right drift speed to the electrons as a function of 
geometry and time (we did not see any correlation between the effects). These 
functions were extracted for our entire data sets and implemented in the 
reconstructions code.

@@

The drift paths are the trajectories which an electron follows in the gas. The 
standard software to calculate drift paths is the MAGBOLTZ program 
~\cite{MAGBOLTZ}. For its calculations, MAGBOLTZ needs to know the detector's 
geometry, the exact composition of the gas mixture and of course the electric 
and the magnetic fields. We used such calculations as a first calibration, but 
as can be seen with the drift speed, conditions are far from stable in the 
chamber. Moreover, because we use of a thin foil as a cathode, the geometry 
knowledge is limited to a millimeter precision. These problems, already 
encountered for the BoNuS RTPC calibration (ref), have motivated the 
acquisition of specific calibration runs. These are taken with lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He). In this process the knowledge of the 
electron kinematic gives us the helium nuclei kinematic and therefore allow us 
to calibrate our drift paths independently of our knowledge of the exact 
conditions in the chamber.

The drift paths' parametrization is possible using identified elastic events 
from our experiment and simulate copy of them in a GEANT4 
simulation~\cite{GEANT4}. Then by comparing the GEANT4 calculated position of 
the helium nuclei to the list of hits effectively measured in the chamber, we 
can reconstruct the electrons drift path in the RTPC. However, because of the 
magnetic field, the drift paths are not linear. So to perform the extraction, 
we make a first assumption of a linear dependence between the radius of 
emission and the TDC of detection and then refine our result. As it happens, 
the curvature is minimal and this process converge already on the second 
iteration. Here are the step of the extraction procedure (we work in a polar 
coordinate system ($R$,$\phi$), $\Delta\phi$ being the difference between the 
$\phi$ of the pad measuring a hit and the $\phi$ of the simulated helium):

\begin{figure}[tb]
\centering
%\hspace{-0.95cm}
\includegraphics[scale=0.37]{fig/TdcR_check_p1_10.png}
\caption{R versus TDC distribution for the simulated hits locating in a single 
$z$ bin. The maximum $R$ equals 60 mm (close to the anode) at minimum TDCs 
(Trigger time is 15), and minimum R equals 30 mm (close to the cathode) at 
maximum TDCs ($\sim75$).}
\label{fig:R_correlation}
\end{figure}

\begin{itemize}

   \item The linear assumption between R and time(TDC) is necessary to link the 
      generated hits from GEANT4 to physical hits. Figure 
      \ref{fig:R_correlation} shows our initial $R$ versus TDC.

   \item For these selected hits, we look at $\Delta\phi$ ($\Delta \phi$ = 
      $\phi_{sim.}$ - $\phi_{hit\_pad}$) versus TDC. We clearly obtain a 
      correlation shown in figure~\ref{fig:1st_pass_delta_phi}. This $\Delta 
      \phi$ change with TDC shows the azimuthal drift of the electrons due to 
      the Lorentz angle.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/TdcPhi_p1_10.png}
         \caption{$\Delta \phi$ versus TDC distribution for the simulated hits 
         locating in one slice centered at 35 mm.}
         \label{fig:1st_pass_delta_phi}
      \end{figure}

   \item Using this first measurement, we can calculate a new correlation 
      between $R$ and TDC. The result, shown in figure~\ref{fig:R_TDC}, is very 
      close to linear explaining the fast convergence of our procedure.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/R_TDC.png}
         \caption{The calculated R versus the time is corrected by the 
         parameters of $\Delta \phi$ from the first pass.}
         \label{fig:R_TDC}
      \end{figure}

@@

   \item We finally parametrize the $\Delta\phi$ distributions with a fit as 
      can be seen in figure~\ref{fig:DELTA_PHI_TDC}, which shows the final 
      $\Delta \phi$ versus TDC with a red line representing our drift paths 
      fit.
\end{itemize}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.37]{fig/FitResult_p2_10.png}
\caption{$\Delta \phi$ versus TDC distribution in the second pass for the slice 
centered at 35 mm longitudinal position along the RTPC. The red line represents 
the final drift paths in this slice.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data set from the beginning of the run period and the 
1.269 GeV data from the end of the run period. As a result, we found very 
similar drift paths and reached the conclusion that whatever changed in the 
setting affected only the drift speed.

@@

The parametrization of the drift speed and the drift paths were carried out 
using only the TDCs, gain calibration goal is to equalize the gains in ADC/MeV 
over the full RTPC. The gain of each pad is the ratio between the actual 
deposited energy and the registered ADC value. We identified three different 
methods to extract such gains. The first is to compare the experimental energy 
deposit to the expected value calculated from Bethe-Block formula (ref). The 
second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by simulated tracks (using the same elastic events than for the drift 
paths calibration). In the third method, the ADCs of each pad is compared to 
the ADCs of the other pads inside each track.

(Detail of 1st method: Nathan)

To perform the second method it requires a very good GEANT4 simulation 
including drift paths, but also drift and amplification spreads of the charges 
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
can compare simulation to data on an event by event basis as in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is defined 
as the ratio of the measured ADCs to the simulated deposited energy.

(Detail of 3rd method: Mohammad)

>>>>

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
Laboratory to track low-energy nuclear recoils for the purpose of measuring
exclusive nuclear channels, such as coherent Deeply Virtual Compton Scattering
and coherent meson production on $^4$He. In such processes, the $^4$He nucleus
remains intact and recoils as a whole in the final state. The CEBAF Large
Acceptance Spectrometer (CLAS) has been used for similar studies on proton
targets, however it cannot track recoils with momenta less than 250 MeV/c. In
2009, we carried out measurements using the CLAS spectrometer supplemented by
the RTPC positioned directly around a gaseous $^4$He target, allowing a detection
threshold as low as $~$100  MeV/c for $^4$He. This article discuses the design,
work principle, calibration methods and performances of the RTPC.
\end{abstract}

@@

Thomas Jefferson National Accelerator Facility (TJNAF), in Newport News, Virginia,
USA, provides high power electron beams of energy up to 6 GeV and 100$\%$ duty cycle
to three experimental Halls (A, B, C) simultaneously. The CLAS spectrometer, located
in Hall-B, is composed of several sub-detectors and two magnets. The spectrometer
was designed to track charged particles of momentum greater than 250 MeV/c as well 
as detecting real photons. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) for the tracking of charged 
    particles~\cite{DCref} (in blue on the figure).
 \item Superconducting toroidal magnet (in yellow) to bend the trajectories 
    of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Cherenvkov Counters (CC) to distinguish electrons from other negative 
    particles, such as pions, at lower energies~\cite{CCref} (in pink).
 \item Scintillation Counters (SC) to identify hadrons by measuring their 
    time of flight (TOF)~\cite{TOFref} (in red).
 \item Electromagnetic Calorimeters (EC) to distinguish electrons from other negative 
    particles at higher energies and to detect neutral particles (neutrons and 
    photons in particular)~\cite{ECref} (in green).
\end{itemize}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates some adaptations
in this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles where the acceptance of the EC is poor. To extend the range,
an additional Inner Calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}.

%M{\o}ller electrons can be
%very problematic at such small angles, so a 5 Tesla solenoid magnet has been
%added around the target as a magnetic shield against this low energy background.
For coherent nuclear processes, the RTPC was developed to extend the tracking
range of CLAS to low energy nuclei and was used during a one month experimental run
in 2009 with a longitudinally polarized electron beam of energy 6.064 GeV incident 
on a gaseous $^{4}$He target.
%, with the goal of measuring asymmetries in Deep Virtual Compton 
%Scattering (DVCS)~\cite{proposal} and exclusive meson production.

@@

\section{CLAS-EG6 RTPC design} \label{sec_design}
The design of CLAS-EG6 RTPC has been chosen according to physics and 
experimental requirements.  At 6 GeV incident electron beam energy, recoiling 
$^{4}$He nuclei from coherent DVCS have an average momentum (per charge) around
125 MeV/c, while the ``standard'' CLAS detects charged particles above a threshold
of 250 MeV/c. The chosen design also has to fit within the available space inside
a 5 Tesla solenoid magnet surrounding the target \cite{Hyon-suk}. Figure
\ref{fig:RTPC2} shows a photograph of the RTPC before being inserted into the solenoid.

The CLAS-EG6 RTPC is a 200~mm long cylinder with a diameter of 150~mm. As charged 
particles traverse the sensitive gas volume, ionized electrons are released and
drift, under the effect of an applied electromagnetic field, outward perpendicularly
to the beam direction. These electrons are then amplified by three layers of GEMs
and detected by a curved readout system on the external shell, as can be seen in
the figure~\ref{fig:RTPC_1_4}. The EG6-RTPC has two detection halves, but, unlike
the BoNuS RTPC, they share a single chamber.  Each half has independent GEM
amplification systems that cover about 80\% of the azimuthal angle.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of CLAS-EG6 RTPC in a plane perpendicular to 
the beam direction.  See text for description of the elements.} 
\label{fig:RTPC_1_4}
\end{figure}

As seen in figure~\ref{fig:RTPC_1_4}, the CLAS-EG6 RTPC has the following 
substructure, from the beam line to the exterior:\\
\begin{itemize}
   \item The target extends along the RTPC's central z-axis, with a diameter of 
      6 mm and enclosed by a 27-$\mu$m-thick Kapton wall.

   \item The first gas gap covers a radial range of 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions induced
      by M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and electronically grounded.                                                                
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode.

   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is around 500~V/cm 
      (???) and perpendicular to the beam. Field cages are installed at both 
      ends of the detector in this region. Progressive voltage depending on 
      radial position is applied on these to maintain a uniform electric field 
      perpendicular to the beam line at the edges of the RTPC.

   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      represents the anode.

   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM system has been mainly chosen for this detector for its flexibility, 
allowing a curved amplification surface. Also, GEMs are  known to create less 
sparks than other amplicifation technologies for gaseous detectors, which
provides a cleaner detection of slow nuclei like $^4$He. The GEMs for the EG6-RTPC
are made from a Kapton insulator layer sandwiched between two 5 $\mu$m copper
layers. The mesh of each GEM layer is chemically etched with 50$\mu$m diameter
holes with double-conical shapes as can be seen in figure~\ref{fig:GEMs}. A
potential difference of 400 V is applied between the two copper layers of each
GEM, creating a very strong electric field leading to high ionization 
from initial electrons and therefore amplification of the signal. Also, a 
potential difference of 150 V is set between the GEMs to push the amplified 
electrons toward the readout pads. The gain of each GEM layer is of order 100, 
making a total gain of about $10^{6}$.\\

\section{Readout System} \label{sec_readout}
The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.

@@

Figure \ref{fig:PADs} shows a schematic drawing of the size and the alignment 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width.  The shift between the rows allows to reduce aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in figure~\ref{fig:PADs}
shows how pads are grouped to 16 channels pre-amplifiers.  A Time to Digital
Converter (TDC) with 114~ns samples is used to measure the time taken by
the electrons to drift from the ionization point to the readout board.  Signal
amplitude is is measured with an Analog-to-Digital-Converter (ADC) without
specific normalization.

Mention ALTRO chip and reference.

Sparcification.  3 consecutive ADC samples above a threshold of ?, and the three
samples before and after.

@@

{\bf Don't we want to have description and quality of elastic event selection in here somewhere?}

@@

We measure the drift speed using tracks detected in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but geometrical symmetry insures they always travel the same 
distance. By identifying the maximum TDC measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

We first measure the drift speed along the 200 mm RTPC's length to take into 
account variations in the electric and magnetic field, some of which are known 
(see figure \ref{fig:B_MAP}) some might be due to slight shifts in the 
geometry. We verified this by looking at the TDC profile generated in the 
chamber by all our tracks (see figure \ref{fig:TDC_profile}). The time profile 
of registered hits clearly show the dropping edge expected from the geometrical 
considerations previously mentioned. We define a value named $TDC_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram, and this value is inversely proportional to the drift speed. The
result in figure \ref{fig:RunNumber_61551_TDCmax_Zslice} shows the dependence of
the drift speed on position along the beam line axis in the chamber.

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.35]{fig/B_MAP.png}
   \caption[]{Schematic drawing of the solenoid magnetic field vectors inside the 
   RTPC. The $y$-axis is the radial distance from the center of the target, and the
   $x$-axis is the longitudinal position along the RTPC (with its center at $z=0$).} 
\label{fig:B_MAP}
\end{figure}

@@

Due to the non perfect experimental conditions including variations in the 
gas mixture, the drift speed might also change during the month long experimental run.  
Figure \ref{fig:Drift_run_number_1} shows the $TDC_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation over time and accounted for these changes in the drift speed in the
track reconstruction. Dealing with such situations is 
possible by adapting the drift speed as a function of 
geometry and time (we did not see any correlation between the effects). These 
functions were extracted for our entire data set and implemented in the 
reconstructions code.

@@

The drift paths are the trajectories the electrons follows in the gas. The 
standard software to calculate drift paths is the MAGBOLTZ program 
~\cite{MAGBOLTZ}, which requires knowledge of the detector's 
geometry, the exact composition of the gas mixture, and of course the electric 
and the magnetic fields. We used such calculations as a first calibration, but, 
as can be seen with the drift speed, conditions are far from stable in the 
chamber. Moreover, because we use a thin foil as a cathode, the geometry 
knowledge is limited to millimeter precision. These problems, already 
encountered for the BoNuS RTPC calibration (ref), motivated the 
acquisition of specific calibration runs. These are taken with a lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He). In this process, measurement of only the 
electron kinematics gives the Helium nucleus's kinematic and therefore allows 
to calibrate the drift paths independently of our knowledge of the exact 
conditions in the chamber.

The drift paths' parametrization is possible using identified elastic events 
from our experiment and a simulated copy of them in a GEANT4 
simulation~\cite{GEANT4}. Then by comparing the calculated GEANT4 position of 
the Helium nuclei to the corresponding hits measured in the chamber, we 
can reconstruct the electrons drift path in the RTPC. However, because of the 
magnetic field, the drift paths are not linear. So to perform the extraction, 
we make a first approximation with a linear dependence between the radius of 
emission and the TDC of detection, and then refine our result. As it happens, 
the curvature is minimal and this process converge already on the second 
iteration.

The steps of the extraction procedure are as follows, where we work
in a polar coordinate system ($R$,$\phi$).  The azimuthal difference between
the measured and simulated hit position is $\Delta\phi$.

\begin{figure}[tb]
\centering
%\hspace{-0.95cm}
\includegraphics[scale=0.37]{fig/TdcR_check_p1_10.png}
\caption{$R$ versus TDC distribution for the simulated hits in a single 
$z$-bin. The maximum $R$ is 60 mm (close to the anode) at minimum TDCs 
(the trigger time is 15), and the minimum $R$ is 30 mm (close to the cathode) at 
maximum TDCs ($\sim75$).}
\label{fig:R_correlation}
\end{figure}

\begin{enumerate}

   \item The linear assumption between R and time (TDC) is necessary to link the 
      generated hits from GEANT4 to physical hits. Figure 
      \ref{fig:R_correlation} shows our initial $R$ versus TDC.

   \item For these selected hits, we look at $\Delta\phi$ ($\phi_{sim.}-\phi_{hit\_pad}$)
     versus TDC. We clearly obtain a 
      correlation shown in figure~\ref{fig:1st_pass_delta_phi}. The relationship
      between $\Delta\phi$ and TDC encodes the azimuthal drift of the electrons due to 
      the Lorentz angle.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/TdcPhi_p1_10.png}
         \caption{$\Delta \phi$ versus TDC distribution for the simulated hits 
         with $z$ near 35 mm.}
         \label{fig:1st_pass_delta_phi}
      \end{figure}

   \item Using this first measurement, we can calculate a new correlation 
      between $R$ and TDC. The result, shown in figure~\ref{fig:R_TDC}, is very 
      close to linear and explains the fast convergence of our procedure.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/R_TDC.png}
         \caption{The calculated $R$ versus the time is corrected by the 
         parameters of $\Delta \phi$ from the first pass.}
         \label{fig:R_TDC}
      \end{figure}

@@

   \item We finally parametrize the $\Delta\phi$ distributions with a fit as 
      can be seen in figure~\ref{fig:DELTA_PHI_TDC}, which shows the final 
      $\Delta \phi$ versus TDC with a black line representing our drift paths 
      fit.
\end{enumerate}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.37]{fig/FitResult_p2_10.png}
\caption{$\Delta \phi$ versus TDC distribution in the second pass for tracks
near 35 mm longitudinal position along the RTPC. The black line represents 
the final drift paths in this slice.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data from the beginning of the run period and the 
1.269 GeV data from the end of the run period. We found very similar drift paths
for the two data sets and concluded that any changes in the system only
significantly affected the drift speed.

@@

The parametrization of the drift speed and the drift paths were carried out 
using only the TDCs, gain calibration goal is to equalize the gains in ADC/MeV 
over the full RTPC. The gain of each pad is the ratio between the actual 
deposited energy and the registered ADC value. We implemented three different 
methods to extract such gains, with varying degrees of success.

The first method is based on the measured average $\small{\frac{dE}{dX}}$
and the expected value based on the Bethe-Bloch formula (PDG REF).  The total energy
deposition for each $^4$He track is calculated as the sum of all ADC hits
attributed to the track.  The path length of that energy deposition is
calculated as the distance along the reconstructed trajectory between the first
and last hits on the track and corrected to account for regions
of bad channels along the track.  The measured average energy loss is then the
ratio of deposited energy and path length.  The Bethe-Bloch formula is
evaluated at the track's reconstructed momentum, and the ratio with the measured
$\small{\frac{dE}{dX}}$ is taken as a gain scaling factor.  This average gain is
attributed equally to each readout channel contributing to the track, and each
channel's gain is averaged over many tracks.  This gain calibration method is
inherently iterative, and, while it does result in measured $\small{\frac{dE}{dX}}$
much closer to the expected Bethe-Bloch curve than prior to calibration, the
results were not very satisfactory.

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by simulated tracks (using the same elastic events than for the drift 
paths calibration).  This requires a very good GEANT4 simulation 
including drift paths, but also drift and amplification spreads of the charges 
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the data acquisition (DAQ) features 
that can lead to cutting out hits. After setting the simulation properly, we 
can compare simulation to data on an event by event basis as in 
figure~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is defined 
as the ratio of the measured ADCs to the simulated deposited energy.

In the third method, the ADC value of each pad is compared to 
the ADCs of the other pads inside each track.
(Detail of 3rd method: Mohammad)

Do you want to show $log(dE/dX)$?
